Relict deciduous oaks show high resilience to drought despite the effect of land-use legacies

### 2018-11-23 man/figures/bai_cronos.Rmd
update figures

##### 
**Figure 4.** Basal Area Increment (BAI) chronologies of *Q. pyrenaica* for northern population (SJ; *green*) and southern ones: low-elevation (CA-Low; *pink*) and high-elevation (CA-High, *purple*) sites. Shading areas correspond to standard error of the mean. Number of series are displayed in the upper plot. We only show years replicated with # series > 5. Linear trend since 1975 is shown for southern high-elevation site (CA-High) .AREGLAR ESTO
##### 
```{r, warning=FALSE}
cro_sites <- read.csv(here("data/cronos_medias", "cronos_sites.csv"), header = TRUE, sep = ',')

>>>>

**Figure 4.** Basal Area Increment (BAI) chronologies of *Q. pyrenaica* for northern population (SJ; *green*) and southern ones: low-elevation (CA-Low; *pink*) and high-elevation (CA-High, *purple*) sites. Shading areas correspond to standard error of the mean. Number of series are displayed in the upper plot. We only show years replicated with # series > 5. Linear trend since 1975 is shown for southern high-elevation site (CA-High) .AREGLAR ESTO
##### 
```{r, warning=FALSE, echo=FALSE}
cro_sites <- read.csv(here("data/cronos_medias", "cronos_sites.csv"), header = TRUE, sep = ',')

### 2018-11-23 man/figures/browning.Rmd
update figures

```{r load_pkg}
library("tidyverse")
library("here")
library("scales")
library("viridis") # colours 
library("ggthemes") # colours 
```

#####

@@

```{r data}
df <- read.csv(file=here::here('/out/anomalies/evi' ,'percen_browing_greening.csv'), header=TRUE)

>>>>

```{r pkg}
library("tidyverse")
library("here")
library("scales")
library("viridis") # colours 
library("ggthemes") # colours 
```

@@

```{r}
df <- read.csv(file=here::here('/out/anomalies/evi' ,'percen_browing_greening.csv'), header=TRUE)

### 2018-11-23 man/figures/corr_plot_years.Rmd
update figures

---
output: 
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    highlight: null
    reference_docx: ../../templates/template_pagebreak.docx
bibliography: ../../refs/references.bib
csl: ../../refs/ecology.csl
editor_options: 
  chunk_output_type: console
---

@@

```{r, echo=FALSE}
show_text <- FALSE
```
##### 
**Appendix S6.** Correlation between indices of resilience (*Rt*, resistance; *Rc*, recovery; *Rs*, Resilience) using periods of several lengths (2, 3 and 4 years after a drought). Top plots (a, b and c) showing the resilience indices of greenness (EVI) to drought; and bottom plots (d, e, f) the resilience indices of tree-growth (BAI) to drought.

>>>>

---
output: 
  word_document:
---

@@

```{r, echo=FALSE}
show_text <- FALSE
```
##### 
**Appendix S5.** Correlation between indices of resilience (*Rt*, resistance; *Rc*, recovery; *Rs*, Resilience) using periods of several lengths (2, 3 and 4 years after a drought). Top plots (a, b and c) showing the resilience indices of greenness (EVI) to drought; and bottom plots (d, e, f) the resilience indices of tree-growth (BAI) to drought.

### 2018-11-23 man/figures/correla_boot_sites.Rmd
update figures

##### 
**Figure S7.** Correlation among site chronologies (CA-High, CA-Low and SJ) in different time-domains after pre-filtering the time-series with increasing size of the moving-average window (1 to 40 years). Each site chronology was smoothed using centred moving averages with different window sizes (1 to 40 years), and then Pearson's correlation coefficient between the each pair chronologies were calculated. Significance was tested using 1000 boostrap replicates and with 95 % confidence intervals built using the R packgae `boot` [@Canty2016].

>>>>

**Figure S6.** Correlation among site chronologies (CA-High, CA-Low and SJ) in different time-domains after pre-filtering the time-series with increasing size of the moving-average window (1 to 40 years). Each site chronology was smoothed using centred moving averages with different window sizes (1 to 40 years), and then Pearson's correlation coefficient between the each pair chronologies were calculated. Significance was tested using 1000 boostrap replicates and with 95 % confidence intervals built using the R packgae `boot` [@Canty2016].

### 2018-11-23 man/figures/cumulative_prec.Rmd
update figures

#####

@@

```{r}

yeah <- presn %>% filter(csum_mean < (meandf - 1*sddf)) %>% 
  mutate(pos = csum_mean - 50) %>%  as.data.frame()

prec_hydrol_sn <- ggplot(presn, aes(x=hyear, y=csum_mean)) + 
  geom_hline(yintercept = meandf) + 
  geom_hline(yintercept = meandf + 2*sddf, color ='blue', linetype = 'dashed') +
  geom_hline(yintercept = meandf - 2*sddf, color ='red', linetype = 'dashed') +
  geom_hline(yintercept = meandf + 1*sddf, color ='blue', linetype = 'dotted') +
  geom_hline(yintercept = meandf  - 1*sddf, color ='red', linetype = 'dotted') +
  geom_errorbar(aes(ymin = csum_mean - csum_se, ymax= csum_mean + csum_se), width=.1, colour='gray') +
  geom_point(size=2) + theme_bw() + 
  geom_line(size=.25, colour = 'black') + 
  scale_x_continuous(breaks=seq(1950,2015, by=5)) +
  scale_y_continuous(breaks=seq(0,1500, by=250)) + 
  geom_text_repel(
    data = yeah,
    aes(label = hyear), 
    size = 2.5, 
    # position = position_dodge(1), 
    alpha = .5,
    colour = "black",
    segment.size = 0.1, 
    segment.alpha = 0.5,
    point.padding=unit(.7,'lines'),
    nudge_y = -yeah$csum_mean,
    direction = 'x'
    ) +
  theme(panel.grid = element_blank(),
        axis.title = element_text(size = 9),
        axis.text.y = element_text(size = 8)) +
  ylab('Cumulative Precipitation (mm)') + xlab('Year') +
  expand_limits(y=c(0,2000)) 
```

```{r}
p <- prec_hydrol_sn + annotation_custom(ggplotGrob(cp), xmin = 1990, xmax = Inf, ymin=1050)
```

>>>>

@@

```{r plot_prec}
yeah <- presn %>% filter(csum_mean < (meandf - 1*sddf)) %>% 
  mutate(pos = csum_mean - 50) %>%  as.data.frame()

prec_hydrol_sn <- ggplot(presn, aes(x=hyear, y=csum_mean)) + 
  geom_hline(yintercept = meandf) + 
  geom_hline(yintercept = meandf + 2*sddf, color ='blue', linetype = 'dashed') +
  geom_hline(yintercept = meandf - 2*sddf, color ='red', linetype = 'dashed') +
  geom_hline(yintercept = meandf + 1*sddf, color ='blue', linetype = 'dotted') +
  geom_hline(yintercept = meandf  - 1*sddf, color ='red', linetype = 'dotted') +
  geom_errorbar(aes(ymin = csum_mean - csum_se, ymax= csum_mean + csum_se), width=.1, colour='gray') +
  geom_point(size=2) + theme_bw() + 
  geom_line(size=.25, colour = 'black') + 
  scale_x_continuous(breaks=seq(1950,2015, by=5)) +
  scale_y_continuous(breaks=seq(0,1500, by=250)) + 
  geom_text_repel(
    data = yeah,
    aes(label = hyear), 
    size = 2.5, 
    alpha = .5,
    colour = "black",
    segment.size = 0.1, 
    segment.alpha = 0.5,
    point.padding=unit(.7,'lines'),
    nudge_y = -yeah$csum_mean,
    direction = 'x'
    ) +
  theme(panel.grid = element_blank(),
        axis.title = element_text(size = 9),
        axis.text.y = element_text(size = 8)) +
  ylab('Cumulative Precipitation (mm)') + xlab('Year') +
  expand_limits(y=c(0,2000)) 
```

```{r plot_all}
p <- prec_hydrol_sn + annotation_custom(ggplotGrob(cp), xmin = 1990, xmax = Inf, ymin=1050)
```

### 2018-11-23 man/figures/gc_perc_trees.Rmd
update figures

---
output: 
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    highlight: null
    reference_docx: ../../templates/template_pagebreak.docx
bibliography: ../../refs/references.bib
csl: ../../refs/ecology.csl
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE, dpi = 200)
```

```{r}
library("tidyverse")
library("here")
library("dplyr")
library("ggrepel")
library("scales")
library("grid")
library("gridExtra")
library("patchwork")
library("stringr")
library("pander")
library("lemon")
library("magrittr")
```

```{r, echo=FALSE}
show_text <- FALSE
```

##### 
**Figure 8**. Percentage of *Q. pyrenaica* trees affected by GC > 50 % by site. *Black* line shows number of trees (rigth-axis). Data for number of trees > 2 is shown.

```{r}
gc_treesMed <- read.csv(here("data/disturbance", "disturbance_trees_gc_treshold_median.csv"), header = TRUE, sep = ',')
gc_treesMed50 <- read.csv(here("data/disturbance", "disturbance_trees_gc_treshold_median50.csv"), header = TRUE, sep = ',')

@@

pgc_median50 <- gc_treesMed50 %>% 
  filter(total_tree > 2) %>% 
  ggplot() +
  geom_bar(aes(x=year, y=per_pos), fill=colorpositivo, colour=colorpositivo, 
           stat = "identity", position = "dodge") + 
  geom_bar(aes(x=year, y=(-1)*per_neg), fill=colornegativo, colour=colornegativo, 
           stat = "identity", position = "dodge") +
  geom_line(aes(x=year, y=total_tree), colour = 'black') + 
  theme_bw() + ylab('% trees') +
  geom_hline(yintercept=0, alpha=.4) +
  geom_hline(yintercept=c(50, -50), linetype=2, alpha=.4) +
  scale_x_continuous(breaks = xrange) +
  scale_y_continuous(limits = c(-100, 100), breaks = seq(-100, 100, by = 50),
                     sec.axis = sec_axis(~ ., name = '# trees', breaks = c(0,15,30))) +
  facet_rep_wrap(~site_sorted, nrow = 3, 
                 repeat.tick.labels = TRUE, labeller = as_labeller(label_site)) +
  theme(panel.grid = element_blank(), 
        strip.background = element_blank(), 
        strip.text = element_blank(),
        axis.text.y = element_text(size=10),
        axis.text.x = element_text(size=9),
        axis.title = element_text(size = 12))

pgc_median50
```

>>>>

---
output: 
  word_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```

```{r load_pkg}
library("tidyverse")
library("here")
library("scales")

library("grid")
library("gridExtra")
library("patchwork")
library("ggcorrplot")
library("stringr")
library("pander")
library("lemon")
```

**Figure S4.** Percentage of *Q. pyrenaica* trees affected by GC > 50 % by site. *Black* line shows number of trees (rigth-axis). Data for number of trees > 2 is shown.

```{r data}
gc_treesMed <- read.csv(here("data/disturbance", "disturbance_trees_gc_treshold_median.csv"), header = TRUE, sep = ',')
gc_treesMed50 <- read.csv(here("data/disturbance", "disturbance_trees_gc_treshold_median50.csv"), header = TRUE, sep = ',')
```

```{r plotGC}

@@

pgc_median50 <- gc_treesMed50 %>% 
  filter(total_tree > 2) %>% 
  ggplot() +
  geom_bar(aes(x=year, y=per_pos), fill=colorpositivo, colour=colorpositivo, 
           stat = "identity", position = "dodge") + 
  geom_bar(aes(x=year, y=(-1)*per_neg), fill=colornegativo, colour=colornegativo, 
           stat = "identity", position = "dodge") +
  geom_line(aes(x=year, y=total_tree), colour = 'black', size = 0.4) + 
  theme_bw() + ylab('% of trees') + xlab("Year") + 
  geom_hline(yintercept=0, alpha=.4) +
  geom_hline(yintercept=c(50, -50), linetype=2, alpha=.4) +
  scale_x_continuous(breaks = xrange) +
  scale_y_continuous(limits = c(-100, 100), breaks = seq(-100, 100, by = 50),
                     sec.axis = sec_axis(~ ., name = '# trees', breaks = c(0,15,30))) +
  facet_rep_wrap(~site_sorted, nrow = 3, 
                 repeat.tick.labels = TRUE, labeller = as_labeller(label_site)) +
  theme(panel.grid = element_blank(), 
        strip.background = element_blank(), 
        strip.text = element_blank(),
        axis.text.y.left = element_text(size=6),
        axis.text.y.right  = element_text(size=5),
        axis.text.x = element_text(size=6),
        axis.title = element_text(size = 9))

```

```{r plot_out, fig.height=4.5, dpi = 300, out.width = 0.8} 
pgc_median50
```

### 2018-11-23 man/figures/plot_spei12.Rmd
update figures

---
output: 
  word_document:
    reference_docx: ../../templates/template_pagebreak.docx
bibliography: ../../refs/references.bib
csl: ../../refs/ecology.csl
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE, dpi = 200)
```

```{r}
library("tidyverse")
library("here")
library("dplyr")
library("ggrepel")
library("scales")
library("stringr")
library("viridis")
library("ggthemes")
```

```{r, echo=FALSE}
show_text <- FALSE
```

##### 
**Appendix S2**. Drought severity in the Sierra Nevada for the 1950-2016 period based on the Standardised Precipitation-Evapotranspiration Index (SPEI). Data from Global SPEI database (http://spei.csic.es/database.html). We obtanied the SPEI data for a 12 month scale and for all 0.5º grid cells covering Sierra Nevada.

```{r, echo=FALSE}
# Get file names
myfiles <- list.files(path = here::here("/data_raw/spei_all/"), pattern = "\\.csv$")

@@

```{r, fig.height=5}
p <- spei %>% 
  ggplot(aes(x=date, y=spei, fill=sign)) +  
  geom_bar(stat = "identity") + 
  scale_fill_manual(values = c("pos" = "#696969", "neg" = "brown4")) +
  theme_bw() + 
  geom_hline(yintercept = 0, colour = 'gray') +
  labs(x = '', y ='SPEI (12-month scale)') +
  scale_x_date(date_breaks = "5 year", date_labels = "%Y", 
               limits=as.Date(c("1949-01-01", "2015-12-01"))) +
  theme(strip.background = element_rect(fill = "white"),
        legend.position = "none", 
        panel.grid = element_blank(), 
        axis.text.y = element_text(size=10),
        axis.text.x = element_text(size=9),
        axis.title.y = element_text(size = 12))

p
```

>>>>

---
output: 
  word_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```

```{r load_pkg}
library("tidyverse")
library("here")
```

```{r data}
# Get file names
myfiles <- list.files(path = here::here("/data_raw/spei_all/"), pattern = "\\.csv$")

@@

```{r plot_spei}
plot_spei <- spei %>%   
  ggplot(aes(x=date, y=spei, fill = sign)) + 
  geom_bar(stat = "identity") + 
  scale_fill_manual(values = c("pos" = "#696969", "neg" = "brown4")) +
  theme_bw() + 
  geom_hline(yintercept = 0, colour = 'gray') +
  labs(x = '', y ='SPEI (12-month scale)') +
  scale_x_date(date_breaks = "5 year", date_labels = "%Y", 
               limits=as.Date(c("1949-01-01", "2015-12-01"))) +
  theme(strip.background = element_rect(fill = "white"),
        legend.position = "none", 
        panel.grid = element_blank(), 
        axis.text.y = element_text(size=10),
        axis.text.x = element_text(size=9),
        axis.title.y = element_text(size = 12))

```

**Appendix S2**. Drought severity in the Sierra Nevada for the 1950-2016 period based on the Standardised Precipitation-Evapotranspiration Index (SPEI). Data from Global SPEI database (http://spei.csic.es/database.html). We obtanied the SPEI data for a 12 month scale and for all 0.5º grid cells covering Sierra Nevada.

```{r plot_out, fig.width=6, dpi = 300, out.width = 0.8} 
plot_spei
```

### 2018-11-23 man/figures/residual_chronos.Rmd
update figures

---
output: 
  word_document:
    fig_caption: yes
    fig_height: 6
    fig_width: 6
    highlight: null
    reference_docx: ../../templates/template_pagebreak.docx
bibliography: ../../refs/references.bib
csl: ../../refs/ecology.csl
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE, dpi = 200)
```

```{r}
library("tidyverse")
library("here")
library("dplyr")
library("ggrepel")
library("scales")

@@

```{r, echo=FALSE}
show_text <- FALSE
```

##### 
**Figure 5.** Residual tree-ring chronologies obtained for the *Q. pyrenaica* sites. Dashed red lines indicate the start of the reliable period (EPS > 0.85). Dotted black lines showing the severe drought years identified in our climatic data (see Table S3).

```{r}
cro_rwi <- read.csv(here("data/cronos_medias", "cronos_sites_rwi.csv"), header = TRUE, sep = ',')
year_cutoffs <- read.csv(here("data/cronos_medias", "cronos_sites_rwi_year_cutoffs.csv"), header = TRUE, sep = ',')
year_sampledepth <- read.csv(here("data/cronos_medias", "cronos_sites_rwi_year_sampledepth.csv"), header = TRUE, sep = ',')

@@

```{r}
# Mayores y menores crecimientos 
# top10  
top10pos50 <- cro_rwi %>% filter(year > 1950) %>% arrange_(~desc(res)) %>% group_by_(~site) %>%  slice(1:10)
top10neg50 <- cro_rwi %>% filter(year > 1950) %>% arrange_(~res) %>% group_by_(~site) %>%  slice(1:10)
top10pos <- cro_rwi %>% arrange_(~desc(res)) %>% group_by_(~site) %>%  slice(1:10)
top10neg <- cro_rwi %>% arrange_(~res) %>% group_by_(~site) %>%  slice(1:10)
```

```{r}
label_site <- c("0_sj" = "SJ",
                "1_caH" = "CA-High",
                "2_caL" = "CA-Low")

@@

# 
plot_rwi_crono <- ggplot(cro_rwi, aes(x=year, y=res)) +
  theme_bw() + 
  theme(panel.grid = element_blank(), 
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face="bold", size=12), 
        # strip.text = element_blank(), 
        axis.text = element_text(face="bold", size=8), 
        axis.title = element_text(face="bold", size=12)) + 
  geom_hline(yintercept = 1, colour = "grey") + 
  geom_line(size=.8) + scale_y_continuous(breaks = seq(0.4,1.8, 0.2)) +
  facet_wrap(~site_sorted, scales = "free_x", ncol = 1,
             labeller = as_labeller(label_site)) + 
  ylab("Ring width index") + xlab("Year") + 
  geom_vline(aes(xintercept = year), data = year_cutoffs,  linetype = "dashed", colour='red') + 
  # geom_vline(aes(xintercept = year), data = year_sampledepth,  linetype = "dashed", colour='blue') +
  geom_vline(xintercept = dys, linetype = "dotted") + 
  # geom_vline(xintercept = dys, linetype = "dotted") + 
  scale_x_continuous(limits = c(1819, 2016), 
                     breaks= seq(1820, 2020, by=20))

plot_rwi_crono  
```

>>>>

---
output: 
  word_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
```

```{r load_pkg}
library("tidyverse")
library("here")
library("scales")

@@

```{r data}
cro_rwi <- read.csv(here("data/cronos_medias", "cronos_sites_rwi.csv"), header = TRUE, sep = ',')
year_cutoffs <- read.csv(here("data/cronos_medias", "cronos_sites_rwi_year_cutoffs.csv"), header = TRUE, sep = ',')
year_sampledepth <- read.csv(here("data/cronos_medias", "cronos_sites_rwi_year_sampledepth.csv"), header = TRUE, sep = ',')

@@

```{r data_aux}
# Mayores y menores crecimientos 
# top10  
top10pos50 <- cro_rwi %>% filter(year > 1950) %>% arrange_(~desc(res)) %>% group_by_(~site) %>%  slice(1:10)
top10neg50 <- cro_rwi %>% filter(year > 1950) %>% arrange_(~res) %>% group_by_(~site) %>%  slice(1:10)
top10pos <- cro_rwi %>% arrange_(~desc(res)) %>% group_by_(~site) %>%  slice(1:10)
top10neg <- cro_rwi %>% arrange_(~res) %>% group_by_(~site) %>%  slice(1:10)
```

```{r plot_rwi}
label_site <- c("0_sj" = "SJ",
                "1_caH" = "CA-High",
                "2_caL" = "CA-Low")

@@

# 
plot_rwi_crono <- ggplot(cro_rwi, aes(x=year, y=res)) +
  theme_bw() + 
  theme(panel.grid = element_blank(), 
        strip.background = element_rect(fill = "white"),
        strip.text = element_text(face="bold", size=9), 
        # strip.text = element_blank(), 
        axis.text = element_text(size=8), 
        axis.title = element_text(face="bold", size=9)) + 
  geom_hline(yintercept = 1, colour = "grey") + 
  geom_line(size=.8) + scale_y_continuous(breaks = seq(0.4,1.8, 0.2)) +
  facet_wrap(~site_sorted, scales = "free_x", ncol = 1,
             labeller = as_labeller(label_site)) + 
  ylab("Ring width index") + xlab("Year") + 
  geom_vline(aes(xintercept = year), data = year_cutoffs,  linetype = "dashed", colour='red') + 
  # geom_vline(aes(xintercept = year), data = year_sampledepth,  linetype = "dashed", colour='blue') +
  geom_vline(xintercept = dys, linetype = "dotted") + 
  # geom_vline(xintercept = dys, linetype = "dotted") + 
  scale_x_continuous(limits = c(1819, 2016), 
                     breaks= seq(1820, 2020, by=20))

```

**Figure S3.** Residual tree-ring chronologies obtained for the *Q. pyrenaica* sites. Dashed red lines indicate the start of the reliable period (EPS > 0.85). Dotted black lines showing the severe drought years identified in our climatic data (see Table S3).

```{r plot_out, fig.height=4.5, dpi = 300, out.width = 0.8} 
plot_rwi_crono  
```
